### 2011-04-07 extended_catalog/paper/paper.tex
more paper

\documentclass[12pt]{article}

\title{Search for Spatially Extended Sources Using Two Years of Flight
Data}

\usepackage{amsmath}

\begin{document}

\maketitle

\abstract{
We present a new method for analyzing spatially 
extended sources with the Fermi Large Area
Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.
We provide a series of Monte Carlo studies
that were done to validate this tool.
We then apply this tool to test all sources in
the 2FGL Two Year Catalog for extension
and report on the discovery of new extended sources.\cite{2FGL}
}

\section{Introduction}

% * Resolving spatially extended Fermi sources is important for 
%    * studying supernova remnants, 
%    * pulsar wind nebulae, 
%    * galaxies,  
%    * potentially dark matter or new physics.
% * We searched all 2FGL catalog sources
% * Systematic test for a certain kind of (~ small) regularly shaped extended

\section{Analysis Method}

%   * Difficulty of studying spatially extended sources
%   * Pointlike, great features
%   * modifying it to fit spatially extended sources.
%      * semi-analytic convolution formula
%   * importance of fitting position + extension + spectrum simultaneously to apply statistical significance
%   * Monte Carlo Validation
%

Performing morphological studies of sources in the GeV energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent point spread function and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXX^\circ$
at 100MeV to $XXXXX^\circ$ at 100GeV and so the higher energy photons are
significantly more important for studying the shape of sources. One could
ideally look just at the highest energy photons, but the trade off is that
for typical sources many more source events are expected at low energy.
For this reason, it is especially important to use as much of the data
as possible.

The second primary difficult in studying spatially extended sources is
systematic uncertainties in the diffuse emission. Systematic errors are
notoriously difficult and will be discussed more later, but the important
point to make here is that one of the most important diagnostic tools for
studying systematic errors is being able to iterate the analysis while
varying different parameters such as what goes into the diffuse model.

A new analysis tool has been developed to address these unique
requirements for studying spatially extended sources with the Fermi
LAT. The overall approach for this analysis is to perform a maximum
likelihood analysis where the Poisson likelihood for observing the
measured counts is computed given an assumed sky model. The parameters
of the sky model are then fit by maximizing the log of the likelihood.
To study the extension of a source, one can assume a particular shape
for an extended source (e.g. a disk or Gaussian), and fit the spatial
parameters of that source.

In principle, the entire task can be done using the standard science
tools package {\em gtlike}\cite{Science-Tools-gtlike}. But {\em gtlike} is
only able to fit the spectral parameters of the source, so the extension
fitting would have to be done as an external loop. Typically, the run time
for a binned source analysis using {\em gtlike} is several hours so this loop
would be especially time consuming. In principle, parallel computing could
ease this burdon, but parallel fitting algorithms are rather untested
and complex. Furthermore, the machinery required to do something like
this would necessarily be rather complex. What has typically been done by
the Fermi team to study extended sources is to fix the assumed position
of the extended source, either by our knowledge of a source from other
wavelenghts, or by fitting it as a point source. Then, a profile of the
likelihood as a function of extension is developed using {\em gtlike}
and the maximum of this profile. This approach is not optimal because it
needs not be the case that the best fit position of a point source would
be the same as the best fit position of an extended source. Furthermore,
by not fully maximizing the likelihood, this method is not as significant to
extension, especially of faint sources. And finally, since this approach
is so computationally intensive, no large scale Monte Carlo effort has been
launched to validate the method.

Instead, an alternate approach was developed for performing
a morphological analysis. An alternate maximum likelihood fitting
package called {\em pointlike} has been developed by the Fermi
collaboration and we have expanded its functionality to fully support
analyzing spatially extended sources. {\em pointlike} is most thoroughly
described in Matthew Kerr's Ph.D Thesis\cite{Matthew_kerr_phd_thesis}
but the important aspects will be summarized below. What differentiates
{\em pointlike} most significantly from {\em gtlike} is its emphasis
on speed by introducing approximations into the calculation of the
likelihood function while at the same time trying to reduce any
numerical errors that may be introduced.  Like binned {\em gtlike},
{\em pointlike} bins the sky both in position and energy.  Unlike {\em
gtlike}, pointlike relies on the {\em healpix} representation of spatial
bins\cite{healpix_paper} and scales the size of each spatial bin to be
smaller than the point spread function. This is convenient because it no
significant information is lost due to binning at all energies and it is
always a good approximation that the model predicted counts are uniform
across a spatial bin. Furthermore, {\em pointlike} uses a sparse matrix
representation of the spatial bins where only the bins with counts in
them are looped over. The only downside to this is that the integral of
the model predicted counts must be independently calculated.  Effectivly,
{\em pointlike} efficiently interpolates from a totally binned analysis
at low energy (where the resolution is poor and there are many counts)
to an unbinned analysis at high energy (where the resolution is very good
but there are few counts). This approach provides significant savings
in time without introducing large numerical error.

% PseudoHypothesis

\section{Validation}

One can define a parameter $\text{TS}_\text{ext}$ as twice the increase
in log likelihood going from fitting a source with a point hypothesis to fitting
it as an extended hypothesis. So $\text{TS}_\text{ext}=2\times(LL_\text{ext}-LL_\text{ps})$.
In going from fitting a point source to fitting a simple extended source such as
a disk, one degree of freedom, the source width, is added. According to
Wilk's Theorem, the distribution of $\text{TS}_\text{ext}$ for no signal (a true
point source) should follow a $\chi^2$ distribution with one degree of freedom\cite{Wilks_Theorem}.
On the other hand, since the null hypothesis exists on the edge of parameter space,
and so so Wilk's Theorem does not hold\cite{Warnings about Wilk's Thorem}.
One might expect the actual distribution to be a half $\chi^2$ distribution with a delta
function at 0. This was the distribution found to be true for source finding of
gamma ray sources.\cite{Mattox_et_All_Paper}. A Monte Carlo study has been performed to
show how well we can believe the statistical significance.

\section{Extension search method}

% * Plots we make ...

\section{Results}

\section{Interpretation}

\end{document}