### 2011-08-28 extended_catalog/collaboration_meeting_presentation/collaboration_meeting_presentation.tex
collaboration meeting presentation

\documentclass[12pt]{beamer}
\usetheme[navbar=false, bkgimage=false, shadow=true]{Fermi}
\usepackage{amsmath}
\usepackage{xspace}

\newcommand{\fermi}{\textit{Fermi}\xspace}

\title{Search for Spatially Extended \fermi-LAT Sources Using Two Years of Flight
Data}
%\subtitle{\ldots}

\author{Joshua Lande, Markus Ackermann, Stefan Funk}
\institute{SLAC/Stanford}
\email{joshualande@gmail.com}
\date{August 25, 2011}

\begin{document}

\fermititle

\begin{frame}{Overview}
  \begin{itemize}
    \item Category II Paper
    \item Contact Authors: J. Lande, 
    \item Full author list being finalized
    \item 
  \end{itemize}
\end{frame}

\end{document}