### 2011-08-28 extended_catalog/lande_ext_collab_august_2011/introduction.tex
better

\fermititle

\begin{frame}{Overview}
  \begin{itemize}
    \item Category II Paper
    \item Contact Authors: J. Lande, M. Ackermann, S. Funk
    \item Full author list being finalized
    \item Internal Referees: Marianne Lemoine-Goumard and Johann Cohen-Tanugi
    \item Target Journal: ApJ
    \item Status (something about being submitted to internal referees XXXX)
  \end{itemize}
\end{frame}

\begin{frame}{Paper Outline}
  \begin{itemize}
    \item Description/validation of a new method (\pointlike) for analyzing extended sources.
    \item  Calculation of the LAT's sensitivity to spatially extended sources
    \item  Presentation of a new search for spatially extended sources:
      \begin{itemize}
        \item reanalyzing the extension of the 12 extended sources in 2FGL
        \item testing AGN from 2LAC for extension to validate the analysis
        \item presenting on the discovery/interpretation of several new extended sources not in 2FGL.
      \end{itemize}
  \end{itemize}
\end{frame}