### 2011-10-06 resume/v1/lande_resume.tex
adding resume to svn

% Resume taken from: http://linux.dsplabs.com.au/resume-writing-example-latex-template-linux-curriculum-vitae-professional-cv-layout-format-text-p54/

\documentclass[margin,12pt,line]{resume}

\begin{document}
\name{\Large Joshua Lande - Curriculum Vitae}

\begin{resume}

    \section{Contact\\Information}

    Kavli Institute for Particle Astrophysics and Cosmology \hfill (650) 427-0102       \\
    Stanford Linear Accelerator Center                      \hfill jlande@stanford.edu  \\
    2575 Sand Hill Road, Menlo Park, CA-94025, US                                       \\

    \section{Education}

    \textbf{Stanford University}, Stanford, CA \\
    \textsl{Doctor of Philosophy} \hfill \textbf{ September, 2008 -- present}\vspace{-3mm}\\\vspace{-1mm}%
    \begin{itemize}
        \item Expected graduation date: May, 2012
        \item Advisors:  Dr. Stefan Funk
    \end{itemize}

    \textsl{Marlboro College, BS in Physics} \hfill \textbf{September, 2004 -- May, 2008}\vspace{-3mm}\\\vspace{-1mm}%

  \section{Talks and Posters}

      Lande, J., ``Search for Spatially Extended Sources Using Two Years of Flight Data,'' 
      \textsl{2011 Fermi Symposium}, \hfill \textbf{May 2011.}

      Lande, J., ``Analysis of the spatial extent of sources in the Fermi-LAT 1 year catalog,'' 
      \textsl{2009 Fermi Symposium}, \hfill \textbf{December 2009.}

￼￼Lande, J., ``Analysis of Spatially Extended Sources with the Fermi-LAT,''
      \textsl{TeV Particle Astrophysics 2009}, \hfill \textbf{June 2009.}

    \section{Distinctions}

    The Buck Turner Prize, \textsl{Marlboro College}                  \hfill \textbf{2008} \\
    The Ernest Coleman Award for Scholarship and Citizenship, \textsl{SULI Internship} \hfill \textbf{2007} \\

    \section{Professional\\Experience}

    \textbf{Stanford University}, Stanford, CA \\

    \textsl{Teaching Assistant}, Physics 110 \hfill \textbf{2011}\\
    Teaching assistant for undergraduate course in classical mechanics.

    \textsl{Teaching Assistant}, Physics 113 \hfill \textbf{2010}\\
    Teaching assistant for undergraduate course in computational
    methods in physics using MATLAB.

    \textsl{Teaching Assistant}, Physics 23 \hfill \textbf{2009}\\
    Teaching assistant for undergraduate course in introductory
    physics.

    \textsl{SULI Internship at SLAC}, Menlo Park, CA, \hfill \textbf{Summer, 2007}\\
      SULI Internship studying x-ray powder diffraction. 
      Adviser: Dr. Samuel Webb.

    \textsl{REU Internship at UC Davis} Davis, CA, \hfill \textbf{Summer, 2006}\\
        Physics REU studying gravitational lensing. Adviser: Dr. Tony Tyson.

    \textsl{SEAP Internship at Analytic Services} Arlington, VA, \hfill \textbf{Summer, 2004 \& 2005}\\
        George Washington University Science \& Engineering
        Apprentice Program studying space weather.
        Adviser: Dr. Ronald Turner.

\end{resume}
\end{document}