### 2011-12-06 presentations/extended_source_search/lande_ext_collab_august_2011/introduction.tex
moving

\fermititle

\begin{frame}{Overview}
  \begin{itemize}
    \item Category II Paper
    \item Contact Authors: J. Lande, M. Ackermann, S. Funk
    \item Internal Referees: Marianne Lemoine-Goumard and Johann Cohen-Tanugi
    \item Target Journal: ApJ
    \item Currently submitted to internal referees
    \item Feedback welcome
  \end{itemize}
\end{frame}

\begin{frame}{Paper Outline}
  \begin{itemize}
    \item Description of a new method (\texttt{pointlike}) for analyzing extended sources.
    \item Monte Carlo calculation of false detection rate for extended sources.
    \item Calculation of the LAT's detection threshold 
      to spatially extended sources
    \item Presentation of a new search for spatially extended sources:
      \begin{itemize}
        \item Reanalyzing extension of 12 extended sources in 2FGL
        \item Test AGN from 2LAC for extension to validate the analysis
        \item Present 9 extended sources not in 2FGL
      \end{itemize}
  \end{itemize}
\end{frame}