### 2012-01-19 pwncatalog/PWNCAT2/paper/sections/analysis_of_the_fermi_lat_data.tex

\section{Analysis of the \fermi-LAT data}

Analysis goes here\ldots