### 2012-01-19 pwncatalog/PWNCAT2/paper/sections/off_peak_phase_selection.tex

\section{Off-peak Phase Selection}

Off-peak goes here\ldots