### 2012-01-21 pwncatalog/PWNCAT2/paper/sections/results.tex
new tables

\section{Results}

Results goes here\ldots

\input{tables/off_pulse_all_energy.tex}

\input{tables/off_pulse_each_energy.tex}

\input{tables/off_pulse_cutoff_test.tex}