### 2012-01-21 pwncatalog/PWNCAT2/paper/sections/results.tex
fixing tables

\section{Results}

Results goes here\ldots

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablecaption{All Energy spectral fit for the 52 LAT-detected Pulsars}
\input{tables/off_pulse_all_energy.tex}
\tablecomments{lala}
\end{deluxetable}

\begin{deluxetable}{ccccccc}
\tabletypesize{\scriptsize}
\tablecaption{Energy bin spectral fit for the 52 LAT-detected Pulsars}
\input{tables/off_pulse_each_energy.tex}
\end{deluxetable}

\begin{deluxetable}{lrrrr}
\tabletypesize{\scriptsize}
\tablecaption{Spectral fitting of pulsar wind nebula candidates with low energy component.}
\input{tables/off_pulse_cutoff_test.tex}
\end{deluxetable}