### 2012-01-21 pwncatalog/PWNCAT2/paper/sections/results.tex
new appendix

\section{Results}

Results goes here\ldots

\tabref{off_pulse_all_energy} shows the results of the all energy analysis of the off-peak emission for each pulsar.

\tabref{off_pulse_each_energy} shows the results of the analysis in separate energy bins of each pulsar.

\tabref{off_pulse_cutoff_test} shows the results of the cutoff test for pulsars with significant low-energy emission.

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablecaption{All Energy spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:off_pulse_all_energy}
}
\input{tables/off_pulse_all_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{ccccccc}
\tabletypesize{\scriptsize}
\tablecaption{Energy bin spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:off_pulse_each_energy}
}
\input{tables/off_pulse_each_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{lrrrr}
\tabletypesize{\scriptsize}
\tablecaption{Spectral fitting of pulsar wind nebula candidates with low energy component
\label{tab:off_pulse_cutoff_test}
}
\input{tables/off_pulse_cutoff_test.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}