### 2012-01-25 pwncatalog/PWNCAT2/paper/sections/off_peak_phase_selection.tex
update tables + new figure

\section{Off-peak Phase Selection}

Off-peak goes here\ldots

\figref{off_pulse_select} shows the off pulse selection for some pulsars\dots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{off_pulse_select_bw.eps}
  \else
  \plotone{off_pulse_select_color.eps}
  \fi
  \caption{Off pulse selection for some pulsars\dots}
  \label{fig:off_pulse_select}
\end{figure}