### 2012-01-26 pwncatalog/PWNCAT2/paper/sections/off_peak_phase_selection.tex
updated text

\section{Off-peak Phase Selection}

Off-peak goes here\ldots

\figref{off_peak_select} shows the off peak selection for some pulsars\dots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{off_peak_select_bw.eps}
  \else
  \plotone{off_peak_select_color.eps}
  \fi
  \caption{Off peak selection for some pulsars\dots}
  \label{fig:off_peak_select}
\end{figure}

THe off peak phase range is defined in \tabref{off_peak_phase_range}.

\begin{deluxetable}{lcccc}
  \tabletypesize{\scriptsize}
  \tablecaption{Observatories, off-pulse definitions and distances of the \todo[inline]{How many pulsars?} \lat-detected pulsars
  \label{tab:off_peak_phase_range}
  }
  \input{tables/off_peak_phase_range}
  \tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}