### 2012-01-26 pwncatalog/PWNCAT2/paper/sections/results.tex

\section{Results}

Results goes here\ldots

\tabref{off_peak_all_energy} shows the results of the all energy analysis of the off-peak emission for each pulsar.

\tabref{off_peak_each_energy} shows the results of the analysis in separate energy bins of each pulsar.

\tabref{off_peak_cutoff_test} shows the results of the cutoff test for pulsars with significant low-energy emission.

\begin{deluxetable}{lccc}
\tabletypesize{\scriptsize}
\tablecaption{All Energy spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:off_peak_all_energy}
}
\input{tables/off_peak_all_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{lcccccc}
\tabletypesize{\scriptsize}
\tablecaption{Energy bin spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:off_peak_each_energy}
}
\input{tables/off_peak_each_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{lcccc}
\tabletypesize{\scriptsize}
\tablecaption{Spectral fitting of pulsar wind nebula candidates with low energy component
\label{tab:off_peak_cutoff_test}
}
\input{tables/off_peak_cutoff_test.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\todo[inline]{Add table on PWN Variability}

\figref{cutoff_test} shows the cutoff test\ldots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{cutoff_test_bw.eps}
  \else
  \plotone{cutoff_test_color.eps}
  \fi
  \caption{Cutoff test for some pulsars\dots}
  \label{fig:cutoff_test}
\end{figure}