### 2012-01-28 pwncatalog/PWNCAT2/paper/tables/off_peak_localization.tex

\tablehead{\colhead{PSR} & \colhead{$\ts_\text{point}$} & \colhead{GLON} & \colhead{GLAT} & \colhead{Pos Err} & \colhead{Offset} & \colhead{\tsext} & \colhead{Extension}\\ \colhead{ } & \colhead{ } & \colhead{(deg)} & \colhead{(deg)} & \colhead{ } & \colhead{(deg)} & \colhead{ } & \colhead{(deg)}}
\startdata
J0007+7303 & 82.8 & 119.66 & 10.46 & None & 0.00 & 10.6 & $<2.62$ \\
J0034$-$0534 & 42.4 & 111.53 & -68.03 & 0.06 & 0.04 & 0.0 & $<0.12$ \\
J0218+4232 & 34.7 & 139.56 & -17.53 & 0.08 & 0.05 & 0.0 & $<0.15$ \\
J0340+4130 & 25.1 & 153.81 & -11.00 & 0.06 & 0.04 & 0.0 & $<0.13$ \\
J0534+2200 & 4959.1 & 184.55 & -5.79 & 0.01 & 0.01 & 0.0 & $<0.02$ \\
J0633+1746 & 2842.4 & 195.12 & 4.22 & 0.02 & 0.05 & 3.3 & $<0.09$ \\
J0835$-$4510 & 304.7 & 263.46 & -3.15 & 0.08 & 0.37 & 295.3 & $0.73 \pm 0.06$ \\
\enddata