### 2012-02-05 pwncatalog/PWNCAT2/paper/sections/results.tex
new extension ul code

\section{Results}

Results goes here\ldots

\tabref{all_energy} shows the results of the all energy analysis of the off-peak emission for each pulsar.

\tabref{each_energy} shows the results of the analysis in separate energy bins of each pulsar.

\tabref{cutoff_test} shows the results of the cutoff test for pulsars with significant low-energy emission.

\begin{deluxetable}{l*{5}c}
\tabletypesize{\scriptsize}
\tablecaption{All Energy spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:all_energy}
}
\input{tables/all_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\clearpage
\begin{deluxetable}{l*{9}c}
%\tabletypesize{\tiny}
\setlength{\tabcolsep}{0.04in}
\tabletypesize{\scriptsize}
\tablewidth{0pt}
\rotate
\tablecaption{Energy bin spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:each_energy}
}
\input{tables/each_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{l*{6}c}
\tabletypesize{\scriptsize}
\tablecaption{Spectral fitting of pulsar wind nebula candidates with low energy component
\label{tab:cutoff_test}
}
\input{tables/cutoff_test.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

The localization results are in \tabref{localization}

\begin{deluxetable}{l*{7}c}
\tabletypesize{\scriptsize}
\tablecaption{Localization and extension fitting results
\label{tab:localization}
}
\input{tables/localization.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\figref{cutoff_test} shows the cutoff test\ldots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{cutoff_test_bw.eps}
  \else
  \plotone{cutoff_test_color.eps}
  \fi
  \caption{Cutoff test for some pulsars\dots}
  \label{fig:cutoff_test}
\end{figure}

\figref{variability} shows the cutoff test\ldots

\begin{deluxetable}{l*{2}c}
\tabletypesize{\scriptsize}
\tablecaption{Variability test
\label{tab:variability}
}
\input{tables/variability.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}