### 2012-05-15 pwncatalog/PWNCAT2/paper/pwncat2.tex
update

\documentclass[12pt,preprint]{aastex}

\bibliographystyle{apj}

%#For adding line numbers:
\usepackage{lineno}
\linenumbers

%\usepackage{rotating}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}

% Change the default indent for itemize/enumerate lists
\usepackage{enumitem}
\setlist{leftmargin=2cm}

% Note, hyperref has to come after other packages!
\usepackage{hyperref}
\usepackage{todonotes}

\usepackage{color}

% LaTeX stylings
\input{style/style.tex}

\graphicspath{{./figures/}}

\begin{document}

\title{PWNCAT2}
\shorttitle{PWNCAT2}

\keywords{
Catalogs;
Fermi Gamma-ray Space Telescope; 
Gamma rays: observations; 
pulsar wind nebula
}

\listoftodos

%\ifdefined\bwfigures
%  Figures are in black and white
%\else
%  Figures are in color
%\fi

\section{Unpulsed magnetospheric emission, and PWN searches}

\input{subsections/timing_analysis.tex}

\input{subsections/off_peak_phase_selection.tex}

\input{subsections/analysis_of_the_fermi_lat_data.tex}

\input{subsections/results.tex}

\input{subsections/discussion.tex}

\bibliography{pwncat2}

\appendix

\input{sections/extension_upper_limits.tex}

\end{document}