### 2012-05-15 pwncatalog/PWNCAT2/paper/subsections/results.tex
new

\section{Results}

First, we tested the sources to see if they
were spatially extended. The localization results are in \tabref{localization}.

\begin{deluxetable}{l*{7}c}
\tabletypesize{\scriptsize}
\tablecaption{Localization and extension fitting results
\label{tab:localization}
}
\input{tables/localization.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

Next, we performed a spectral analysis over all energy using the best
fit morphology. \tabref{all_energy} shows the results of the all energy
analysis of the off-peak emission for each pulsar.

Next, we fit a powerlaw independently in each energy
bin. \tabref{each_energy} shows the results of the analysis in separate
energy bins of each pulsar.

Finally, we tested sources to see which were
variable. \tabref{cutoff_test} shows the results of the cutoff test for
pulsars with significant low-energy emission.

\begin{deluxetable}{l*{5}c}
\tabletypesize{\scriptsize}
\tablecaption{All Energy spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:all_energy}
}
\input{tables/all_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\clearpage
\begin{deluxetable}{l*{9}c}
%\tabletypesize{\tiny}
\setlength{\tabcolsep}{0.04in}
\tabletypesize{\scriptsize}
\tablewidth{0pt}
\rotate
\tablecaption{Energy bin spectral fit for the \todo[inline]{How many pulsars?} \lat-detected Pulsars
\label{tab:each_energy}
}
\input{tables/each_energy.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{deluxetable}{l*{6}c}
\tabletypesize{\scriptsize}
\tablecaption{Spectral fitting of pulsar wind nebula candidates with low energy component
\label{tab:cutoff_test}
}
\input{tables/cutoff_test.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\figref{cutoff_test} shows the cutoff test\ldots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{cutoff_test_bw.eps}
  \else
  \plotone{cutoff_test_color.eps}
  \fi
  \caption{Cutoff test for some pulsars\dots}
  \label{fig:cutoff_test}
\end{figure}

\figref{variability} shows the variability test for each source
candidate. The distribution of \tsvar is plotted in

\begin{deluxetable}{l*{2}c}
\tabletypesize{\scriptsize}
\tablecaption{Variability test
\label{tab:variability}
}
\input{tables/variability.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\begin{figure}
  \ifdefined\bwfigures
  \plotone{variability_bw.eps}
  \else
  \plotone{variability_color.eps}
  \fi
  \caption{Distribution of \tsvar for each source candidate. \todo[inline]{Disclaimer about crab not being included}.}
  \label{fig:variability}
\end{figure}