### 2012-06-06 pwncatalog/PWNCAT2/paper/subsections/results.tex

\subsection{Results}

First, we tested the sources to see if they
were spatially extended. The localization results are in \tabref{offpeak}.

\begin{deluxetable}{l*{7}c}
\tabletypesize{\scriptsize}
\tablecaption{Off-Peak Spatial and Spectral Results
\label{tab:offpeak}
}
\input{tables/off_peak_spatial_spectral.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

Next, we performed a spectral analysis over all energy using the best
fit morphology. XXX shows the results of the all energy
analysis of the off-peak emission for each pulsar.

Next, we fit a powerlaw independently in each energy
bin. XXX shows the results of the analysis in separate
energy bins of each pulsar.

Finally, we tested sources to see which were
variable. XXX shows the results of the cutoff test for
pulsars with significant low-energy emission.

\figref{cutoff_test} shows the cutoff test\ldots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{cutoff_test_bw.eps}
  \else
  \plotone{cutoff_test_color.eps}
  \fi
  \caption{Cutoff test for some pulsars\dots}
  \label{fig:cutoff_test}
\end{figure}

We tested all sources for variability.

Special cases:
\begin{itemize}
  \item Crab has a funny spectrum, we have (so far) fixed it to
  \item \velax
\end{itemize}