### 2012-06-07 pwncatalog/PWNCAT2/paper/subsections/off_peak_phase_selection.tex
update

\subsection{Off-peak Phase Selection}
\label{subsec:off_peak_defintion}

To study the off-peak emission of \lat-detected pulsars, we first
developed a new method for defining the off-peak emission.
The primary constraint for this method was that it was systematic,
computationally efficient, and model independent.

The method we developed procedes by deconstructing a pulsar phaseogram
into a Baysian blocks using the implementation described in \cite{Jackson_Bayesian_Blocks_2003}.

\todo[inline]{This is just a placeholder. Hopefully Andrea can fill this section
in with the real details.}

\todo[inline]{How to implement Baysian blocks on periodic data.}

\figref{off_peak_select} shows the off peak selection for some pulsars\dots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{off_peak_select_bw.eps}
  \else
  \plotone{off_peak_select_color.eps}
  \fi
  \caption{Off peak selection for some pulsars. 
  \todo[inline]{This is just a placeholder figure. 
  Hopefull Andrea or somebody can update the plot
  once the real algorithm is finished.}}
  \label{fig:off_peak_select}
\end{figure}