### 2012-07-19 pwncatalog/PWNCAT2/paper/subsections/offpeak_results.tex

\subsection{Results}

After analyzing the off-peak emission using the pipeline described
in \subsecref{off_peak_analysis}, we consider any emission in the
off-peak region to be magnetospheric in nature if the emsision is
not significantly-extended and has a significantly-cutoff spectrum.
We consider the emission to originate in the pulsar wind if it is
spatially extended or had a hard spectral index.
If the source is point-like and had a soft spectrum, or if it 
is spatially extended but the extension is biased by modeling
of the background, we are unable to determine the origin of the emission.

A summay of the reuslts of the pipepline can be found in \tabref{offpeak}.
It includes off off-peak phase ranges selected using the method described
in \subsecref{off_peak_defintion}.
This table includes \ts, \tsext, and \tscut
for the significantly detected pulsars (with $\ts>25$).
This table also includes the best fit flux and spectral index for these pulsars assuming
the best hypothesis (either point-like, spatially-extended, or expontentially cutoff).
For the sources that are significantly cutoff, the spectral index refers to the index
from the fit of a cutoff spectral model and the final column includes the cutoff energy derived from the fit.

\begin{deluxetable}{l*{7}c}
\tabletypesize{\scriptsize}
\tablecaption{Off-Peak Spatial and Spectral Results
\label{tab:offpeak}
}
\input{tables/off_peak_spatial_spectral.tex}
\tablecomments{\todo[inline]{Put table comments}}
\end{deluxetable}

\figref{cutoff_test} shows the cutoff test\ldots

\begin{figure}
  \ifdefined\bwfigures
  \plotone{cutoff_test_bw.eps}
  \else
  \plotone{cutoff_test_color.eps}
  \fi
  \caption{Cutoff test for some pulsars\dots}
  \label{fig:cutoff_test}
\end{figure}

Consistent with \citep{LAT_Collaboration_Crab_Flare_2011}, we found
the Crab nebula to be highly variable with \tsvar=XXXX.  Besides that,
we found no significantly variable off-peak emission.  The results of
the variability test are contained in the auxiliary information.

The results of the spectral analysis in smaller energy bands described
in section \secref{peak_analysis} are included in the auxiliary
information. In addition, upper limits computed assuming a powerlaw
spectral model and a canonical pulsar spectrum are included in the
auxiliary information.

Results:
\begin{itemize}
  \item List of all XXX detections
  \item For each pulsar, is it a 
    \begin{itemize}
      \item Magnetospheric emission
      \item PWN emission
      \item Region where the fit shows trouble for some reason.
    \end{itemize}
\end{itemize}

In our analysis, we detected the the Vela X PWN (associated with PSRJ0835$-$4510), the Crab Nebula (associated
with PSRJ0534+2200), and MSH 15$-$52 (associated with PSRJ1513-5908).

What to say about PSRJ1023-5746

Sources we detect as clearly magnetospheric

Special cases/problem cases:
\begin{itemize}
  \item PSRJ0101-6422: Very weird spectrum
  \item PSRJ1112-6103: Very close to MSH 11~62 (arXiv:1202.3371v1). Nearby source is 2FGL J1112.1-6040. Significant extension.
  \item When fitting PSRJ1119-6127, description of adding nearby point
  source to represent emission residual from o represent residual from
  PSRJ1112-6103 looking spatially
    extended.
  \item PSRJ1418-6058 and PSRJ1420-6048, very difficult to analyze because both very near eachother.
  \item Something about the complciated Gamma Cygni PSR: PSRJ2021+4026
  \item PSRJ1105-6107: something about adding source to background model that is very nearby. \todo{See what OZLEM is doing}.
  \item PSRJ1648-4611 (what to do about nearby 2FGL soruce and other residual emission. Is this paper relevant: http://arxiv.org/pdf/1111.2043.pdf???)?
  \item What to do about emission from PSR J1023 at low energy???
  \item PSRJ0908-4913: waht to say about far localization?
\end{itemize}

\todo[inline]{Describe how we present the point-like hypothesis for signifciant extended soures if the extension fails for some reason}