### 2012-10-16 papers/tevcat/sections/acknowledgements.tex
16/10/2012 : latest version with discussion and conclusion

The \textit{Fermi} LAT Collaboration acknowledges generous ongoing support
from a number of agencies and institutes that have supported both the
development and the operation of the LAT as well as scientific data analysis.
These include the National Aeronautics and Space Administration and the
Department of Energy in the United States, the Commissariat \`a l'Energie Atomique
and the Centre National de la Recherche Scientifique / Institut National de Physique
Nucl\'eaire et de Physique des Particules in France, the Agenzia Spaziale Italiana
and the Istituto Nazionale di Fisica Nucleare in Italy, the Ministry of Education,
Culture, Sports, Science and Technology (MEXT), High Energy Accelerator Research
Organization (KEK) and Japan Aerospace Exploration Agency (JAXA) in Japan, and
the K.~A.~Wallenberg Foundation, the Swedish Research Council and the
Swedish National Space Board in Sweden.

Additional support for science analysis during the operations phase is
gratefully acknowledged from the Istituto Nazionale di Astrofisica in
Italy and the Centre National d'\'Etudes Spatiales in France.

This research has made use of pywcsgrid2, an open-source plotting package for Python6\footnote{http://leejjoon.github.com/pywcsgrid2/}. The authors acknowledge the use of HEALPix\footnote{\url{http://healpix.jpl.nasa.gov/}} \citep{2005ApJ...622..759G}.

The authors acknowledge the use of the TeV catalog website\footnote{http://tevcat.uchicago.edu/} provided by the university of Chicago.

The authors acknowledge the use of the \emph{Australia Telescope National Facility} pulsar catalog\footnote{http://www.atnf.csiro.au/people/pulsar/psrcat/}.