### 2012-10-16 papers/tevcat/sections/build_the_candidate_list.tex

\section{List of candidates}
As a starting point, we used the catalog of TeV sources provided by the university of Chicago\footnotemark[1] to build our list of candidates. This catalog summarizes the information of all sources detected by the TeV experiments. Looking for PWNe candidates which are Galactic sources, we selected the sources located within $\pm$5$\degr$ in latitude.

The Galactic center being a complex region to study with \emph{Fermi}-LAT, due to the contamination by the large density of sources and by the diffuse emission, we removed all sources within 5$\degr$ of the Galactic center. Therefore, we did not include to our list the three sources HESS~J1745-303, HESS~J1741-302 and G~0.9+0.1 \citep{Mouse, Mouse2}.

TeV sources associated to SNRs will not be discussed in this analysis. These sources will be included in a forthcoming paper which aims at bringing constraints on all known SNRs using the \emph{Fermi}--LAT data \cite{SNRcat}.

Finally we removed from this analysis the Crab Nebula and Vela-X. Both were already studied in details (\cite{2010ApJ...708.1254A}, \cite{2012ApJ...749...26B}), and the second is the object of a new analysis \cite{VelaX}.

The final list of 58 sources studied in this analysis is summarized in Table~\ref{tab:TeV_sources} together with their morphology as seen in TeV.