### 2012-10-18 fermi/tevcat/paper/sections/abstract.tex
Josh's changes to the abstract

Since its launch, the \emph{Fermi} satellite has firmly identified seven pulsar wind nebulae (PWNe) and a large number of PWNe candidates, all powered by young and energetic pulsars. Furthermore, PWNe are the most populous class of sources detected at TeV energys followed by the unidentified sources (UNID). Using 45 months of \emph{Fermi}--LAT data, we studied the 10 \gev to 316 \gev $gamma$-ray emission around the position of 58 TeV sources that are either associated with PWNe or unidentified. For each source, we derived a $\gamma$--ray flux or a flux upper limit (when the TeV source is not detected at GeV energies by the \emph{Fermi}--LAT).

The wealth of multi-wavelength data available and the new results provided by the \emph{Fermi}--LAT provide an extraordinary opportunity to constrain the origin of the $\gamma$-ray emission of the large sample of UNID and the radiative processes taking place in known PWNe.