### 2012-10-30 fermi/radiopsrs/paper/sections/notes.tex
update

\section{notes}

Benefits:
\begin{enumerate}
\item Improved time range (XXX months vs XXX months in 2PC)
\item Improved uppper limit from likelihood test.
\item Better analysis method (search for extended sources, could be more sensitive?)
    Look for hard-index sources (unlike 2PC).
\item For example, Lande et al 2012 discovered Gamma-ray emission
    from HESS J1616-508 which is associated with PSR J1617-5055.
\item  Note: Table 3 and 4 get upper limits assuming cutoff spectrum, not suitable
for PWNe searches. For describing of 2PC flux upper limits, see 
\url{https://confluence.slac.stanford.edu/x/UwDNBw}
\end{enumerate}

Notes/questions:
\begin{enumerate}
  \item Should we used reprocessed data?
\end{enumerate}