### 2012-10-30 fermi/gamma_quiet_psrs/paper/sections/introduction.tex
update

\section{Introduction}

\begin{itemize}
  \item Pulsar Physics
  \item PWNe (high energy IC emission)
  \item PWNe detected at GeV energies
  \begin{itemize}
    \item 2PC Off-peak PWNe search
    \item TeVCat PWNe search
  \end{itemize}
  \item $\gamma$-quiet pulsars
  \item Search for new pulsars. Put some kind of note about how LAT has uniform coverage of all sources, which is a benefit.
  \item Also put note about how we discovered a source HESS J1616 \& HESS J1825 at \gev energies using this method.
\end{itemize}

\citep{LAT_Collaboration_2FGL_2012}.