### 2012-11-08 fermi/gamma_quiet_psrs/paper/sections/introduction.tex
update

\section{Introduction}

The Large Area Telescope (LAT) on board the \fermi Gamma-ray Space Telescop has vastly improved
the field of $\gamma$-ray pulsars.

How many pulsars detected by the LAT.

Most prominently, the LAT Collaboratino rece

\begin{itemize}
  \item Pulsar Physics
  \item Pulsars causing a pulsar wind.
  \item PWNe (high energy IC emission)
  \item PWNe detected at GeV energies
  \begin{itemize}
    \item 2PC Off-peak PWNe search
    \item TeVCat PWNe search
    \item Why we see high Edot pulsars.
  \end{itemize}
  \item Pulsars not seen by the LAT ($\gamma$-quiet pulsars)
  \item Search for new pulsars. Put some kind of note about how LAT has
  uniform coverage of all sources, which is a benefit.
  \item Also put note about how we discovered a source HESS J1616 \&
  HESS J1825 at \gev energies using this method.
\item Put note about how there are 17 pulsars with Edot>1e37 (table 3 in 2PC).
  Of those, 8 has no gamma-ray pulsations. There are 46 with $10^{36}$ < Edot < $10^{37}$.
  19 pulsars with no gamma-ray pulsations (table 4 in 2PC). 
  In total, there are 27 pulsars with $E_dot>10^{36}$ with no pulsations
  (despite thorough search). We search the LAT data for these 27 high Edot pulsars to find
  new PWNe.
\end{itemize}

\citep{LAT_Collaboration_2FGL_2012}.