### 2012-12-03 fermi/gamma_quiet_psrs/paper/sections/method.tex
update

\section{Analysis Method}

\subsection{Pulsar Selection Criteria}

Our selection criteria for our PWN search was that to selected high $\dot E$
pulsars with no $\gamma$-ray pulsations.

We took Table 10 from 2PC [CITATION NEEDED],

We excluded J0537$-$6910 and J0540$-$6919. They are in the LMC region which has been detected.

Exclude 3 regiongs with dedicated publications:
\begin{enumerate}
  \item J1617$-$5055 (HESS J1616-508)
  \item J1838$-$0655 (HESS J1837-069)
  \item J1826$-$1334 (HESS J1825-137)
\end{enumerate}

Don't include pulsars which have LAT-detected pulsars since publication of 2FGL.
Presumably, need to redo analysis in off-peak
\begin{enumerate}
  \item J1824$-$2452A 
  \item J1055$-$6028
\end{enumerate}

Don't include sources in Romain's PWNe search:
\begin{enumerate}
  \item J1301-6305 -> HESS J1303-631
  \item J1813-1749 -> HESS J1813-178
\end{enumerate}

\subsection{LAT Data Preparation}