### 2012-12-04 fermi/gamma_quiet_psrs/paper/sections/method.tex
add

\section{Analysis Method}

\subsection{Pulsar Selection Criteria}

Our selection criteria for our PWN search was that to selected high $\dot E$
pulsars with no $\gamma$-ray pulsations.

We took Table 10 from 2PC [CITATION NEEDED],

We excluded PSR J0537$-$6910 and PSR J0540$-$6919. They are in the LMC region which has been detected.

Exclude 3 regiongs with dedicated publications:
\begin{enumerate}
  \item PSR J1617$-$5055 (HESS J1616-508)
  \item PSR J1838$-$0655 (HESS J1837-069)
  \item PSR J1826$-$1334 (HESS J1825-137)
\end{enumerate}

Don't include pulsars which have LAT-detected pulsars since publication of 2FGL.
Presumably, need to redo analysis in off-peak
\begin{enumerate}
  \item PSR J1824$-$2452A 
  \item PSR J1055$-$6028
\end{enumerate}

Don't include sources in Romain's PWNe search:
\begin{enumerate}
  \item PSR J1301-6305 -> HESS J1303-631 (detected by Romain)
  \item PSR J1813-1749 -> HESS J1813-178 (not detected by Romain)
  \item PSR J1809-1917 -> HESS J1809-193 (not detected by Romain)
  \item PSR J1849-0001 -> HESS J1849-000 (not detected by Romain)
  \item PSR J1831-0952 -> HESS J1831-098 (not detected by Romain)
  \item PSR J1913+1011 -> HESS J1912+101 (not detected by Romain)
  \item PSR J1846-0258 -> HESS J1846-029 (not detected by Romain)
  \item PSR J1856+0245 -> HESS J1857+026 (detected by Romain + his previous dedicated publication)
  \item PSR J1803-2137 -> HESS J1804-216 (significantly detected-http://arxiv.org/pdf/1109.3017v2.pdf)
  \item PSR J1930+1852 -> VERJ1930+188 (not detected by Romain)
  \item PSR J1614-5048 is spatially coincident with HESS J1616-508 (detected in ext. src. search paper).
    Note that this HESS source cannot be energetically powered by
    PSR J1614-5048 (see discussion in ext. src. search paper), but
    the region itself was already studied, so we discard this source.

\end{enumerate}

\subsection{LAT Data Preparation}